---
title: "Health Insurance in the United States"
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bootswatch: lux
    orientation: rows
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(readr)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(dplyr)
library(highcharter)
library(viridisLite)
library(ggplot2)
```

```{r}
knitr::opts_chunk$set(cache = TRUE)

data(uscountygeojson)

thm <- 
  hc_theme(
    colors = c("#1a6ecc", "black", "#90ed7d"),
    chart = list(
      backgroundColor = "transparent",
      style = list(fontFamily = "Helvetica")
    ),
    xAxis = list(
      gridLineWidth = 1
    )
  )

data("USArrests", package = "datasets")
data("usgeojson")

Data = read_csv("https://raw.githubusercontent.com/davidye007/USA_Insurance/main/US_Insurance_Dataset.csv")
Base_data = Data %>%
  filter(InsuranceType != "Total") %>%
  filter(InsuranceType != "Any Coverage") %>%
  mutate(Percent = ifelse(is.na(Percentage), 0, Percentage)) %>%
  mutate(Percent_Decimal = Percent * 0.01) %>%
  mutate(Population_expanded = Population * 1000)

Blue_states = c("Washington", "Oregon", "California", "Minnesota", "Illinois", "New York", "Vermont", "Maine", "Massachussetts", "Rhode Island", "Conneticut", "New Jersey", "Delaware", "Maryland", "District of Columbia", "Hawaii")
Red_states = c("Idaho", "Montana", "Wyoming", "Utah", "North Dakota", "South Dakota", "Nebraska", "Kansas", "Oklahoma", "Texas", "Missouri", "Arkansas", "Louisiana", "Mississippi", "Alabama", "South Carolina", "Tennessee", "Kentucky", "West Virginia", "Alaska")
Not_consistent = c("Nevada", "Arizona", "Colorado", "New Mexico", "Iowa", "Wisconsin", "Michigan", "Indiana", "Ohio", "Pennsylvania", "New Hampshire", "Virginia", "North Carolina", "Georgia", "Florida")

federal_minimum_wage = c("Alabama","Georgia", "Idaho", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Mississippi", "New Hampshire", "North Carolina", "North Dakota", "Oklahoma", "Pennsylvania", "South Carolina", "Tennessee", "Texas", "Utah", "Wisconsin", "Wyoming")
non_indexed = c("Arkansas", "Delaware", "Hawaii", "Illinois", "Maryland", "Massachusetts", "New Mexico", "New York", "Rhode Island", "West Virginia")
national = c("United States")
```

Explore States
================================

Questions to Explore: {.sidebar}
-------------------------------------

1. How does heath insurance enrollment vary across the United States?

2. How did the enrollment rate for different types of insurance change over time in each state?

```{r}
selectInput("year", label = "Year:",
            choices = rev(seq(2008, 2022, 1)[-which(seq(2008, 2022, 1) == 2020)]), 
            selected = 2022)

selectInput("insur_type", label = "Insurance:",
            choices = c("Direct-purchase", "Employer-based", "Medicaid", "Medicare", "Private", "Public", "TRICARE", "Uninsured", "VA_Care"), selected = "Direct-purchase")
```

Row
-------------------------------------

### Insurance across the U.S.

```{r, cache=FALSE}
renderHighchart({
  
  selectedState <- reactiveVal("Washington")  # Reactive value to store the selected state
  ### Will make plot dynamic eventually
  Base_Data_filtered <- Base_data %>% filter(InsuranceType == input$insur_type, Year == input$year, State != "United States")
  
  min_value <- min(Base_Data_filtered$Percentage)  # Calculate the minimum value
  
  
  highchart() %>%
    hc_add_series_map(usgeojson, Base_Data_filtered, name = "Percent of State: ", value = "Percentage", joinBy = c("woename", "State")) %>%
    hc_mapNavigation(enabled = TRUE) %>%
    hc_add_theme(thm) %>%
    hc_title(text = paste(input$insur_type, "Insurance in ", input$year), style = list(color = "black", fontSize = "22px")) %>%
    hc_plotOptions(
      series = list(
        events = list(
          click = JS("function(event) { Shiny.onInputChange('selected_state', event.point.name);}")
        )
      )
    )%>%
    hc_colorAxis(
      min = min_value,  # Set the minimum value for the color scale
      max = NULL,  # You can also set the maximum value if needed
      labels = list(
        format = "{value}%",
        style = list(fontSize = "14px")
      )
    )%>%
    hc_mapNavigation(
      zoomIn = list(
        enabled = TRUE
      ),
      zoomOut = list(
        enabled = TRUE
      )
    )
})
```

### Time-Series Plot of Insurance Type (click on the map to change states)

```{r}
observe({
  print(input$selected_state)  # This will print the selected year in the R console
})

```


```{r, cache=FALSE}
custom_palette <- brewer.pal(9, "Dark2")
renderPlot({
  filtered_data <- Base_data %>% 
    filter(State == ifelse(is.null(input$selected_state), "Washington", input$selected_state))
  
  sel_state <- ifelse(is.null(input$selected_state), "Washington", input$selected_state)
  
  ggplot(filtered_data, aes(x=Year, y=Percent, color = reorder(InsuranceType, desc(Percent))))+
  geom_line(size = 0.9)+
    scale_y_continuous(trans='log10', labels = function(x) paste0(x, "%")) +  # Modify this line to add 'r' to each tick mark on the y-axis
  #scale_y_continuous(trans='log10')+
  #scale_colour_brewer(palette = "Dark2")+
  scale_fill_manual(values = custom_palette)+
  ggtitle(paste("Insurance Enrollment For", sel_state))+
  xlab("Year")+
  ylab("Percentage of State (log10 scale)") +
  labs(color='Insurance Type') +
  theme(plot.title = element_text(size=22, hjust = 0.5, color="black"),
        axis.title.x=element_text(size=16, color="black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title=element_text(size=16, color="black", hjust = 0.5),
        legend.text=element_text(size=14, color="black",),
        legend.key.size=unit(0.4, units = "cm"),
        legend.justification = c("right", "top"),
        legend.background = element_rect(fill = "white"),  # Set legend background to black
        legend.key = element_rect(fill = "white", color = "white"),  # Set legend key background and border color to black
        axis.title.y=element_text(size=16, color="black"),
        panel.grid.major = element_line(color= alpha("black", 1)),  # Set major gridlines to black
        panel.grid.minor = element_line(color=alpha("black", 1)),  # Remove minor gridlines
        panel.background = element_rect(fill="white"),  # Set background color to grey
        plot.background = element_rect(fill = "white",color = "white"),  # Set the entire plot background to black
        text = element_text(family = "Helvetica")
    )
})
```


Explore By Classification
================================
Questions to Explore: {.sidebar}
-------------------------------------
SPACE 

3. Does heath insurance enrollment rate and main insurance type change based on a state's political affiliation?

4. Does heath insurance enrollment rate and main insurance type change based on a state's Minimum Wage Status?

```{r}


selectInput("insur_type_2", label = "Insurance:",
            choices = c("Direct-purchase", "Employer-based", "Medicaid", "Medicare", "Private", "Public", "TRICARE", "Uninsured", "VA_Care"), selected = "Direct-purchase")
```

Row {data-height=1000}
-------------------------------------
### Time-Series Plot By A State's Political Affiliation

```{r, cache=FALSE}

renderPlot({
  df_political_affiliation = Base_data %>%
  filter(InsuranceType == input$insur_type_2) %>%
  mutate(Political_Affiliation = case_when(State %in% Blue_states ~ "Democratic",
                            State %in% Red_states ~ "Republican",
                            State %in% Not_consistent ~ "Inconsistant",
                            TRUE ~ "Not a State")) %>%
  filter(Political_Affiliation != "Not a State")%>%
  group_by(Political_Affiliation, Year, InsuranceType)%>% 
  summarise(average_percent = mean(Percent_Decimal))
  
  ggplot(df_political_affiliation, aes(x=Year, y=average_percent, color =Political_Affiliation)) +
  geom_line(size = 0.9)+
  scale_color_manual(values=c("blue", "black", "red"))+
  labs(color='Political Affilitation') +
  ggtitle(paste(input$insur_type_2, "Insurance Enrollment"))+
  xlab("Year") +
  ylab("Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  expand_limits(y = 0) +  # Set expand argument to force y-axis to start from zero
  # scale_y_continuous(trans='log10')+
  theme(plot.title = element_text(size=22, hjust = 0.5, color="black"),
        axis.title.x=element_text(size=16, color="black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title=element_text(size=16, color="black", hjust = 0.5),
        legend.text=element_text(size=14, color="black",),
        legend.key.size=unit(0.4, units = "cm"),
        legend.position = c(0.8, 0.2),
        legend.background = element_rect(fill = "lightgrey"),  # Set legend background to black
        legend.key = element_rect(fill = "white", color = "white"),  # Set legend key background and border color to black
        axis.title.y=element_text(size=16, color="black"),
        panel.grid.major = element_line(color= alpha("black", 1)),  # Set major gridlines to black
        panel.grid.minor = element_line(color=alpha("black", 1)),  # Remove minor gridlines
        panel.background = element_rect(fill="white"),  # Set background color to grey
        plot.background = element_rect(fill = "white",color = "white"),  # Set the entire plot background to black
        text = element_text(family = "Helvetica")
    )
  
})
```

### Time-Series Plot By A State's Minimum Wage Type

```{r, cache=FALSE}

renderPlot({
  df_Minimum_Wage = Base_data %>%
  filter(InsuranceType == input$insur_type_2) %>%
  mutate(Minimum_Wage = case_when(State %in% non_indexed ~ "Above Federal",
                            State %in% federal_minimum_wage ~ "At or Below Federal",
                            TRUE ~ "Indexed")) %>%
  filter(State != "United States")%>%
  group_by(Minimum_Wage, Year, InsuranceType)%>% 
  summarise(average_percent = mean(Percent_Decimal))  
  
  ggplot(df_Minimum_Wage, aes(x=Year, y=average_percent, color = Minimum_Wage)) +
  geom_line(size = 0.9)+
  scale_color_manual(values=c("green", "red", "blue"))+
  labs(color='Minimum Wage') +
  ggtitle(paste(input$insur_type_2, "Insurance Enrollment"))+
  xlab("Year") +
  ylab("Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  expand_limits(y = 0) +  # Set expand argument to force y-axis to start from zero
  # scale_y_continuous(trans='log10')+
  theme(plot.title = element_text(size=22, hjust = 0.5, color="black"),
        axis.title.x=element_text(size=16, color="black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title=element_text(size=16, color="black", hjust = 0.5),
        legend.text=element_text(size=14, color="black",),
        legend.key.size=unit(0.4, units = "cm"),
        legend.position = c(0.8, 0.2),
        legend.background = element_rect(fill = "lightgrey"),  # Set legend background to black
        legend.key = element_rect(fill = "white", color = "white"),  # Set legend key background and border color to black
        axis.title.y=element_text(size=14, color="black"),
        panel.grid.major = element_line(color= alpha("black", 1)),  # Set major gridlines to black
        panel.grid.minor = element_line(color=alpha("black", 1)),  # Remove minor gridlines
        panel.background = element_rect(fill="white"),  # Set background color to grey
        plot.background = element_rect(fill = "white",color = "white"),  # Set the entire plot background to black
        text = element_text(family = "Helvetica")
    )
  
})
```


Row {.tabset .tabset-fade}
-------------------------------------
### Methodology For Political Affiliation Categoraztion and Display
This graph shows the differences in percentage rates by a state's political affiliation. States are categorized by blue, red or non-consistent based on how that have voted in presidential elections since 2020 (https://www.270towin.com/same-since-electoral-maps/). If they have voted the same political party for the full period of time they will be categorized either red or blue, if they have switched parties in that period they will be categorized as a non consistent. 

The y-axis percentage refers to the percent of a state's population enrolled in a given insurance type. Each percent per year on the graph is the average percentage for all states with a certain Political Affiliation.

### Methodology For Minimum Wage Categoraztion and Display
This graph shows the differences in percentage rates by a state's minimum wage type. States are categorized by at or below, above or indexed minimum wage based on a table from the National Conference of State Legislatures (https://www.ncsl.org/labor-and-employment/state-minimum-wages). If a state is "at or below the minimum wage" then the state legislature matches the federal minimum wage, is below federal minimum wage or doesn't specify a minimum wage. If a state is categorized as "above minimum wage" then the state has increased their minimum wage one or more times since 2009, when the federal minimum wage of 7.25 was enacted. If the state has "indexed minimum wage" then the state adjusts or increases the minimum wage on some set schedule that may or may not correlate with some kind of index/measurement (CPI, cost of living, employment cost index etc. ). Minimum wage can be used as an indicator for average income for each state. This is important because public insurance in the United States is regulated by income level.

The y-axis percentage refers to the percent of a state's population enrolled in a given insurance type. Each percent per year on the graph is the average percentage for all states with a certain minimum wage type.